\vspace{-.5cm} Lights: a generalized joint model for high-dimensional multivariate longitudinal data and censored durations \vspace{.5cm}

### 2021-08-20 paper.tex
Update the tex label of grad-Q

\vspace{.5cm}
Lemma~\ref{lemma:prox-sgl1} bellow states that one can easily compute the proximal operator of the sparse group lasso defined in Algorithm~\ref{alg:ISTA}.
\begin{lemma}
\label{lemma:prox-sgl1}
The proximal operator of the sparse group lasso can be expressed as the composition of the group lasso and the lasso proximal operators, that is
\normalfont
\begin{equation*}
  \text{prox}_{\text{sg} l_1,\tilde \eta, \zeta} = \text{prox}_{\zeta \tilde \eta \sum_{l=1}^L \norm{\cdot}_{2,l}} \circ \text{prox}_{\zeta (1 - \tilde \eta) \norm{\cdot}_1}.
\end{equation*}
\end{lemma}
The two proximal operators on the right hand side of Lemma~\ref{lemma:prox-sgl1} are both well known analytically~\citep{bach2011optimization} and tractable. We used the \texttt{Python} library \texttt{copt} for the implementation of proximal gradient descent~\citep{copt} and we propose in our package an implementation for the proximal operator of the sparse group lasso based on Lemma~\ref{lemma:prox-sgl1}.
A proof of Lemma~\ref{lemma:prox-sgl1} is proposed in Appendix~\ref{sec:proof-lemma-prox-sgl1}.
Hence, we update $\gamma_k^{(w)}$ with the following step
\begin{equation}
  \label{eq:gamma-update}
  \gamma_k^{(w+1)} = \text{ISTA}_{\text{sg} l_1}\big(\gamma_k^{(w)}, Q^{(w)}_{n,k}(\cdot), \tilde \eta, \zeta_{2,k}\big)
\end{equation}
Note that the warmstart initialization is left to the user's choice, as well as the ISTA acceleration~\citep{beck2009fast}, since warmstart can diminish acceleration effectiveness~\citep{tibshirani2010proximal}. The gradient of $Q_{n,k}^{(w)}$ is here given by
\begin{align}
  \label{eq:grad-Q-2}
  &\frac{\partial Q_{n,k}^{(w)}(\gamma_k)}{\partial \gamma_{k,a}^l} = -n^{-1} \sum_{i=1}^n \hat \pi_{ik}^{\theta^{(w)}} \Big[ \delta_i  \hat \E_{\theta^{(w)}}[ \varphi_a(t_i, b_i^l) | t_i, \delta_i, y_i] \\ \nonumber
   &\qquad\qquad - \sum_{j=1}^{J-1} (t_{j+1}^u-t_j^u) \lambda_0^{(w)}(t_j^u) \hat \E_{\theta^{(w)}}[ \varphi_a(t_j, b_i^l) \exp \big\{\gamma_k^\top \Psi_i(t_j^u)\big\} | t_i, \delta_i, y_i] \ind{\{t_j^u < t_i\}} \Big]
\end{align}
for all $a = 1, \ldots, \mathcal{A}$ and $l = 1, \ldots, L$.

@@

Considering the convex minimization problem~\eqref{eq:minimization-xi_k} at a given step $w$,
let us denote $\zeta_{1,k}^1 \leq \zeta_{1,k}^2 \leq \dots \leq \zeta_{1,k}^{\text{max}}$ the randomly chosen candidate values for $\zeta_{1,k}$, such that at $\zeta_{1,k}^{\text{max}}$, all coefficients $\hat \xi_{k, j}$ for all $j = 1, \dots, p$ are exactly zero.
The KKT conditions~\citep{boyd2004convex} claim that
\begin{center}
  $\left\{
      \begin{aligned}
        &\frac{\partial P_{n, k}^{(w)}(\hat \xi_k)}{\partial \xi_{k,j}} =  \zeta_{1,k} \big( (\eta-1)\ \text{sgn}(\hat \xi_{k,j}) - \eta \hat \xi_{k,j} \big)  \quad \ \ \ \forall j \in \hat \cA_k \\
        &\left| \frac{\partial P_{n, k}^{(w)}(\hat \xi_k)}{\partial \xi_{k,j}} \right| < \zeta_{1,k} (1 - \eta) \quad \quad \quad \quad \quad \quad \quad \ \forall j \notin \hat \cA_k
      \end{aligned}
    \right.$,
\end{center}
where $\hat \cA_k = \big\{j = 1, \dots, p : \hat \xi_{k,j} \neq 0 \big\}$ is the active set of the $\hat \xi_k $ estimator, and for all $x \in~\R\setminus~\{0\},\ \text{sgn}(x) = \mathds{1}_{\{x>0\}} - \mathds{1}_{\{x<0\}} $. Then, using~\eqref{eq:grad-xi_k}, one obtains
\begin{equation*}
\hat \xi_{k, j} = 0 \ \Rightarrow \ \left| n^{-1} \sum_{i=1}^n \hat \pi_{ik}^{\theta^{(w)}} \big( 1 - \pi_{\xi_k}(x_i) \big) x_{ij}  \right| < \zeta_{1,k} (1 - \eta)
\end{equation*}
for all $j = 1, \dots, p$.
Hence, we choose the following upper bound for the randomly chosen candidate interval during the cross-validation procedure
\begin{equation*}
\zeta_{1,k}^{\text{max}} = \frac{1}{n(1-\eta)} \underset{j = 1, \dots, p}{\text{max}} \sum_{i=1}^n |x_{ij}|.
\end{equation*}
Similar strategy to automatically compute $\zeta_{2,k}^{\text{max}}$ is not easy since without additionnal hypothesis, the gradients~\ref{eq:grad-beta_k},~\ref{eq:grad-Q-1} and~\ref{eq:grad-Q-2} are not upper bounded by a quantity that depends only on the data and not on the current iteration parameters estimates. We then choose those interval upper bounds empirically.

>>>>

\vspace{.5cm}
Lemma~\ref{lemma:prox-sgl1} bellow states that one can easily compute the proximal operator of the sparse group lasso defined in Algorithm~\ref{alg:ISTA}.
\begin{lemma}
\label{lemma:prox-sgl1}
The proximal operator of the sparse group lasso can be expressed as the composition of the group lasso and the lasso proximal operators, that is
\normalfont
\begin{equation*}
  \text{prox}_{\text{sg} l_1,\tilde \eta, \zeta} = \text{prox}_{\zeta \tilde \eta \sum_{l=1}^L \norm{\cdot}_{2,l}} \circ \text{prox}_{\zeta (1 - \tilde \eta) \norm{\cdot}_1}.
\end{equation*}
\end{lemma}
The two proximal operators on the right hand side of Lemma~\ref{lemma:prox-sgl1} are both well known analytically~\citep{bach2011optimization} and tractable. We used the \texttt{Python} library \texttt{copt} for the implementation of proximal gradient descent~\citep{copt} and we propose in our package an implementation for the proximal operator of the sparse group lasso based on Lemma~\ref{lemma:prox-sgl1}.
A proof of Lemma~\ref{lemma:prox-sgl1} is proposed in Appendix~\ref{sec:proof-lemma-prox-sgl1}.
Hence, we update $\gamma_k^{(w)}$ with the following step
\begin{equation}
  \label{eq:gamma-update}
  \gamma_k^{(w+1)} = \text{ISTA}_{\text{sg} l_1}\big(\gamma_k^{(w)}, Q^{(w)}_{n,k}(\cdot), \tilde \eta, \zeta_{2,k}\big)
\end{equation}
Note that the warmstart initialization is left to the user's choice, as well as the ISTA acceleration~\citep{beck2009fast}, since warmstart can diminish acceleration effectiveness~\citep{tibshirani2010proximal}. The gradient of $Q_{n,k}^{(w)}$ is here given by
\begin{align}
  \label{eq:grad-Q}
  &\frac{\partial Q_{n,k}^{(w)}(\gamma_k)}{\partial \gamma_{k,a}^l} = -n^{-1} \sum_{i=1}^n \hat \pi_{ik}^{\theta^{(w)}} \Big[ \delta_i  \hat \E_{\theta^{(w)}}[ \varphi_a(t_i, b_i^l) | t_i, \delta_i, y_i] \\ \nonumber
   &\qquad\qquad - \sum_{j=1}^{J-1} (t_{j+1}^u-t_j^u) \lambda_0^{(w)}(t_j^u) \hat \E_{\theta^{(w)}}[ \varphi_a(t_j, b_i^l) \exp \big\{\gamma_k^\top \Psi_i(t_j^u)\big\} | t_i, \delta_i, y_i] \ind{\{t_j^u < t_i\}} \Big]
\end{align}
for all $a = 1, \ldots, \mathcal{A}$ and $l = 1, \ldots, L$.

@@

Considering the convex minimization problem~\eqref{eq:minimization-xi_k} at a given step $w$,
let us denote $\zeta_{1,k}^1 \leq \zeta_{1,k}^2 \leq \dots \leq \zeta_{1,k}^{\text{max}}$ the randomly chosen candidate values for $\zeta_{1,k}$, such that at $\zeta_{1,k}^{\text{max}}$, all coefficients $\hat \xi_{k, j}$ for all $j = 1, \dots, p$ are exactly zero.
The KKT conditions~\citep{boyd2004convex} claim that
\begin{center}
  $\left\{
      \begin{aligned}
        &\frac{\partial P_{n, k}^{(w)}(\hat \xi_k)}{\partial \xi_{k,j}} =  \zeta_{1,k} \big( (\eta-1)\ \text{sgn}(\hat \xi_{k,j}) - \eta \hat \xi_{k,j} \big)  \quad \ \ \ \forall j \in \hat \cA_k \\
        &\left| \frac{\partial P_{n, k}^{(w)}(\hat \xi_k)}{\partial \xi_{k,j}} \right| < \zeta_{1,k} (1 - \eta) \quad \quad \quad \quad \quad \quad \quad \ \forall j \notin \hat \cA_k
      \end{aligned}
    \right.$,
\end{center}
where $\hat \cA_k = \big\{j = 1, \dots, p : \hat \xi_{k,j} \neq 0 \big\}$ is the active set of the $\hat \xi_k $ estimator, and for all $x \in~\R\setminus~\{0\},\ \text{sgn}(x) = \mathds{1}_{\{x>0\}} - \mathds{1}_{\{x<0\}} $. Then, using~\eqref{eq:grad-xi_k}, one obtains
\begin{equation*}
\hat \xi_{k, j} = 0 \ \Rightarrow \ \left| n^{-1} \sum_{i=1}^n \hat \pi_{ik}^{\theta^{(w)}} \big( 1 - \pi_{\xi_k}(x_i) \big) x_{ij}  \right| < \zeta_{1,k} (1 - \eta)
\end{equation*}
for all $j = 1, \dots, p$.
Hence, we choose the following upper bound for the randomly chosen candidate interval during the cross-validation procedure
\begin{equation*}
\zeta_{1,k}^{\text{max}} = \frac{1}{n(1-\eta)} \underset{j = 1, \dots, p}{\text{max}} \sum_{i=1}^n |x_{ij}|.
\end{equation*}
Similar strategy to automatically compute $\zeta_{2,k}^{\text{max}}$ is not easy since without additionnal hypothesis, the gradients~\ref{eq:grad-beta_k} and~\ref{eq:grad-Q} are not upper bounded by a quantity that depends only on the data and not on the current iteration parameters estimates. We then choose those interval upper bounds empirically.
